---
title: "Practical Machine Learning Course Project"
author: "LM"
date: "31 stycznia 2016"
output: html_document
---
## Introduction

The goal of this project is to create model that predict  the manner in which they did the exercise using giroscopes and accelerators date.
  
## getting data 
Input data is downloaded from presented url:
```{r,cache=TRUE}
train.input<-
  read.csv(url("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"),
           header=TRUE,
           na.string=c("NA",""))
test.input<- 
  read.csv(url("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"),
           header=TRUE,
           na.string=c("NA",""))
```

## Cleaning data

There are a lot of columns that for many of rows. The first row looks like that is representative. First 7 column are identification columns so are not usefull for algorithm.
```{r,cache=TRUE}
usable.columns<-c()
for(i in 1:ncol(train.input))
{
  if(is.na(train.input[1,i])==FALSE)
  {
    usable.columns<-c(usable.columns,names(train.input[i]))

  }
}

usable.columns<-usable.columns[c(-1:-7)]
train.data<-train.input[,usable.columns]
test.data<-test.input[,usable.columns[usable.columns!="classe"]]

str(train.data)
```

## Model  
To create model, data is divided into two parts:  
1) traning 60%  
2) testing 40%  
```{r,cache=TRUE}
### creating test and train parts
library(caret)
inTrain = createDataPartition(y=train.data$classe, p=0.01, list=FALSE)
train = train.data[inTrain,]
test = train.data[-inTrain,]
```

## exploratory analysis
Polots of first variabiles shows that clasification using random forest will be the best method for this task:
```{r,cache=TRUE}
p1<-ggplot(train,aes(roll_belt,pitch_belt,colour=classe))+stat_density2d() +geom_point(aes(colour = factor(classe)))
p2<-ggplot(train,aes(roll_belt,yaw_belt,colour=classe))+stat_density2d() +geom_point(aes(colour = factor(classe)))
p3<-ggplot(train,aes(pitch_belt,yaw_belt,colour=classe))+stat_density2d() +geom_point(aes(colour = factor(classe)))
p1;p2;p3
pairs(train[1:4],bg = c("red","green3","blue","orange","gray")[unclass(train$classe)],pch=21)
```

##Model traning and testing

The method for calculations is Random Tree. This method has high timecost but has high accurace.
```{r,cache=TRUE}
library(randomForest)
model=train(classe~.,method="rf",data=train)
predictResults=predict(model,test)
mean(predictResults==test$classe)
```
##Conclusion
Random forest model predict classe of excercise with high accurace and can be usefull for predictions.
